Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2016-03-11 Publications/2015-NeuroImage/responseLetter3.tex

\begin{document}
\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}

\section*{Response letter to the manuscript no. NIMG-15-2877R1}

\bigskip
\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-2877R1 submitted to NeuroImage.
We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
Please, find in the following paragraphs our point-by-point answer to the concerns raised by the reviewers,
  along with the description of the corresponding changes in the manuscript.
Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate 
  which reviewer and comment the corresponding change is addressing to.

\bigskip
\bigskip
\subsection*{Reviewer \#1:}
\begin{reviewintro}
In general, I think the authors have done a good job of restructuring the paper to qualify the goals and implementation of the method. I have a few points.
\end{reviewintro}
\begin{reviewpoint}
The authors suggest that diffusion tractography is typically performed by mapping segmentations to the data. This is becoming less true. The HCP for one performs whole brain connectivity analysis. Nevertheless, these frameworks still require registration to structural space for projection to surface mesh representations and/or coaligning data. This may be worth mentioning or clarifying.
\end{reviewpoint}
\reply{%
The reviewer is right: despite new whole-brain connectivity strategies, registration from/to structural space is still required.
We have added this clarification in lines 359--363, in the
  discussion section.
}
\begin{reviewpoint}
I still find figure 2 really difficult to read anything from. It's not easy to read the labels in the coloured text, nor see the different histograms in figure 2 A. This is partly because many of the colours are light very similar i.e. for WM and bst. It does occur to me that not all distributions change that greatly with registration. Maybe stripping back figure A to only include WM, GM and possibly VdGM might help?
\end{reviewpoint}
\reply{%
We have made the following editions in this figure to make it more clear:
\begin{itemize}
\item Changed color palette to the ``Set1'' palette of the ColorBrewer tool\footnote{\url{http://colorbrewer2.org/}}, which has
  darker colors.
\item The darker colors allowed for thiner lines to represent the contours of the joint density plot in panel A, improving
  readability.
\item Better positioned labels in the panel A of this figure, using the Fruchterman-Reingold force-directed algorithm.
\item Matched the fonts of axis labels with the font used in the NeuroImage journal.
\item Rearranged the tissues in panel B to match the hierarchical ordering of tissues.
\item Added annotations to capture the reader's attention and point it to those significant areas of the plot, highlighting the
effects of misregistration in each tissue distribution.
\end{itemize}
}
\begin{reviewpoint}
I find the clarification of how tratography might be unwarped in the response very helpful, but some clarification of the manuscript might be helpful. What do the authors mean by ``the resampling of the DWIs introduced by the correction method can be avoided, either by performing the posterior processing on the native dMRI space''. How does the native dMRI space differ from the DWIs here, or what is meant by posterior processing.
\end{reviewpoint}
\reply{%
We have reformulated that particular sentence (lines 57--59), in a way that
  it avoids mentioning posterior processing.
How this methodological variation would affect posterior processing (analyzing tracks, interpolating FA on them, building the connectivity matrix, etc.) is discussed in the paragraph containing lines 358--359.
Particularly, these lines 358--359 were inserted to address this comment, which is highly related to Comments \ref{com:1}, \ref{com:5} and \ref{com:6}.
}
\newpage
\begin{reviewpoint}
In the appendix I think the B-spline variable (Eqs A.7 A.8)  should be defined in the text again
\end{reviewpoint}
\reply{%
We have defined introduced the definition of $\mathbf{g}_k$ in lines 122--123, right before the
  reference to the mentioned equations.
}

@@

Please clarify that the authors do not consider this to be an EPI distortion correction method, but rather a surface alignment method or similar. This was not clear to me from reading the paper the first time. I assumed EPI distortion correction was a primary goal, but it seems rather that the primary goal is to accurately segment the distorted diffusion images. In fact, the distortions are not corrected in the EPI data.
\end{reviewpoint}
\reply{%
We have explicitly stated this in a new sentence (lines 355--357).

@@

Please discuss the fact that when this method is applied, the tractography is done using data that have not been corrected for EPI distortions.
\end{reviewpoint}
\reply{%
Right after the previous edition, in lines 356--357 we introduce the proposal of this methodological variation.
The discussion and proposal of a future research line is then done in lines 363--364, after some further details added

@@

The ``undistorted'' HCP data is not fully defined in the ``Real datasets'' section. Because this undistortion method serves as the ground truth for the presented experiments, it should be explained briefly how the undistortion was done for the HCP data. One sentence is sufficient, and will allow the reader better understand the results of this paper.
\end{reviewpoint}
\reply{%
We have introduced this description in lines 184--188, along with including the detail of the actual HCP release of the data used in l. 178.

@@

\begin{reviewpoint}
Minor points:

The graphical abstract says subpixel accuracy. I think you mean subvoxel (there are no pixels really in 3D MRI images).

``preliminar DTI'' -> preliminary
\end{reviewpoint}
\reply{%
We have updated the graphical abstract with the term ``subvoxel''.
We have replaced the six appearances of the term ``pixel'' by ``voxel'', including the ``subpixel'' present in line 61.
We also have fixed the misspelling of ``preliminar'' of line 350.
}

### 2016-03-11 Publications/2015-NeuroImage/06-discussion.tex
new figure, MBC comments

The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
Current processing of \gls{dmri} involved in the connectome extraction and other applications
  (such as \gls*{tbss} or surgical planning) require a precise segmentation
  of the anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
We envision \regseg{} to be integrated in diffusion processing pipelines, after a 
  \revcomment[R\#3-C10]{preliminary}
  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
  methods.
Since the structural information is projected into the native space of \gls*{dmri},
  these two processes and the matrix building task can be performed on the unaltered
  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
  can be used to map the tracts into structural space.
\revcomment[R\#3-C5]{%
Even though we apply \regseg{} to the problem of susceptibility distortion,
  \emph{it is not a distortion correction method}, but rather a surface alignment method.
In fact, the distortions are not corrected in the \gls*{epi} data.}
\revcomment[R\#3-C6]{%
Therefore, we propose here to perform the reconstruction and tractography processes in the
  original (distorted) diffusion data.}
\revcomment[R\#1-C1]{%
Depending on the application (like \gls*{tbss}, or connectivity analyses that do not build the
  network matrix upon anatomical information mapped from a reference image), the obtained tracks
  may be geometrically unwarped using the estimation of the deformation found by \regseg{}.
Conversely, when the information necessary in the diffusion analysis is projected from the reference
  space, \regseg{} allows to avoid resampling and/or unwarping.}
\revcomment[R\#3-C6]{%
This methodological variation should be further investigated, to ensure which processing design
  yields the most accurate tractography results.}

>>>>

The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
Current processing of \gls{dmri} involved in the connectome extraction and other applications
  (such as \gls*{tbss} or surgical planning) require a precise segmentation
  of the anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is found commonly in historical datasets.
We envision \regseg{} to be integrated in diffusion processing pipelines, after a 
  \revcomment[R\#3-C10]{preliminary}
  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
  methods.
Since the structural information is projected into the native space of \gls*{dmri},
  these two processes and the matrix building task can be performed on the unaltered
  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
  can be used to map the tracts into structural space.
\revcomment[R\#3-C5]{%
Even though we apply \regseg{} to the problem of susceptibility distortion,
  \emph{it is not a distortion correction method}, but rather a surface alignment method.
In fact, the distortions are not corrected in the \gls*{epi} data.}
\revcomment[R\#3-C6]{%
Therefore, we propose here to perform the reconstruction and tractography processes in the
  original (distorted) diffusion data.}
\revcomment[R\#1-C3]{%
\Regseg{} allows to avoid resampling and/or unwarping because the information necessary in the diffusion analysis is projected
  from the reference space.}
\revcomment[R\#1-C1]{%
Certain applications (like \gls*{tbss}) and methodologies (like building the connectivity matrix
  by clustering the tracks) may not perform on the native (distorted) diffusion space correctly,
  but they still need a mapping to the reference space.
Using \regseg{}, the tracks obtained in native space can be unwarped using the resulting estimation
  of the deformation field.}
\revcomment[R\#3-C6]{%
This methodological variation should be further investigated, to ensure which processing design
  yields the most accurate tractography results.}
